Investigating shape standardization using geometric morphometry of pottery from Iron Age northeastern Taiwan

### 2020-04-10 analysis/paper/paper.Rmd
edit Archaeological background and materials

Ceramics analyzed in this paper come from 40 units (4m by 4m each) sampled from the central, undisturbed area of archaeological excavations at Kiwulan (Figure 1; Figure 2). The chronology of the archaeological deposits at Kiwulan consists of two cultural components, the upper component and the lower component, with a sterile layer in between (Chen, 2007). This paper focuses on the upper component, dated from AD 1350 to AD 1950, covering the late Iron Age and the historical period, which we define as the European occupation in Taiwan in the early 17th century. The Dutch first occupied southern Taiwan in 1624 and then the Spanish occupied northern Taiwan in 1626 (Andrade, 2007). In 1642, the Spanish was expelled by the Dutch in northern Taiwan, who took over their forts at Helping Dau in Keelung, and in Tamsui. Since then, western Taiwan was mostly under Dutch colonial rule until 1662 when the Kingdom of Tungning in Taiwan was founded by Koxinga, a loyalist of the Ming dynasty of China (Andrade, 2007).

>>>>

Ceramics analyzed in this paper come from 40 units (4m by 4m each) sampled from the central, undisturbed area of archaeological excavations at Kiwulan (Figure \@ref(fig:KWL-map); Figure \@ref(fig:KWL-sam-area)). The chronology of the archaeological deposits at Kiwulan consists of two cultural components, the upper component and the lower component, with a sterile layer in between [@Chen2007]. This paper focuses on the upper component, dated from AD 1350 to AD 1850, covering the late Iron Age and the historical period, which we define as the European occupation in Taiwan in the early 17th century. The Dutch first occupied southern Taiwan in 1624 and then the Spanish occupied northern Taiwan in 1626 [@Andrade2007]. In 1642, the Spanish was expelled by the Dutch in northern Taiwan, who took over their forts at Helping Dau in Keelung, and in Tamsui. Since then, western Taiwan was mostly under Dutch colonial rule until 1662 when the Kingdom of Tungning in Taiwan was founded by Koxinga, a loyalist of the Ming dynasty of China [@Andrade2007].
